---
title: "Treemap Visualizations"
output:
  rmarkdown::github_document: default
---

```{r echo=FALSE, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

### Notebook objective
- Explore using treemap visualizations to show hierarchical relationships and distributions

###  Packages
```{r}
required_packages <- c('tidyverse', 'treemap')
for(p in required_packages) {
  ### un-comment the below lines to install packages not installed
  # if(!require(p,character.only = TRUE))
  #       install.packages(p, repos = "http://cran.us.r-project.org")
  library(p,character.only = TRUE)
}
```

### Load example dataset
- data included with base R
- warning population data values are estimates from 1975

```{r}
estimated_state_pop_1975 <- tibble(
      state_full_name = rownames(state.x77),
      # we need to divide pop value by "1000" to put the var on millions scale
      state_population_estimated_1975_millions = as.data.frame(state.x77)$Population/1000
)

state_df <- tibble(state_abb = state.abb,
       state_full_name = state.name,
       region = state.region) %>%
      left_join(estimated_state_pop_1975)
```

### States Treemap Example
```{r treemap1, dev = "png"}
treemap(state_df,
        index=c("region","state_abb"),
        vSize="state_population_estimated_1975_millions",
        vColor="state_population_estimated_1975_millions",
        type="value",
        palette = "Spectral",
        align.labels=list(c("center", "center"), c("left", "top")),
        title="Treemap example 1: 1975 US state population")
```

```{r treemap2, dev = "png"}
treemap(state_df,
        index=c("region","state_abb"),
        vSize="state_population_estimated_1975_millions",
        type="index",
        palette = "Pastel1",
        align.labels=list(c("center", "center"), c("left", "top")),
        title="Treemap example 2: 1975 US state population")
```

### Packages for inactive treemaps
- [d3tree](https://cran.r-project.org/web/packages/d3Tree/d3Tree.pdf)
- [plotly](https://cran.r-project.org/web/packages/plotly/plotly.pdf)

### Takeaways
- Interactive plot likely needed for complex treemap
- Proportionally small observations are hard to see on the plot
- In practice, treemap usefulness will depend on the dataset and should be compared to vs bar chart simplicity
- Pick the visualization that helps the audience get to insights/action fastest

